Query Optimization for Dynamic Imputation

### 2017-03-01 paper/algorithm.tex

\subsection{Imputation operators}
\label{sec:operators}
We introduce two new relational operators to perform imputation: \textit{Impute} ($\mu$) and
\textit{Drop} ($\delta$). Each operator takes arguments $(C, R)$ where $C$ is a set of
attributes and $R$ is a relation. \textit{Impute} uses a machine learning algorithm to
replace all \nullv{} values with non-null values for attributes in $C$ in the relation $R$ (discussed in detail in~\Cref{sec:imputation}).
In our experiments, we instantiate \textit{Impute} with the CE-CART~\cite{burgette2010multiple} algorithm, but we highlight that
our system is agnostic to this choice.
\textit{Drop} simply removes from $R$ all tuples which have a \nullv{} value for some attribute in $C$.
Both operators guarantee that the resulting relation will contain no \nullv{} values for
attributes in $C$.

@@

\begin{definition}
Let $P$ be a set of plans. For $p \in P$, we say $p$ is $\alpha$-bound, if $(\textsc{Loss}(p) - \min_{p' \in P}(\textsc{Loss}(p')) \leq \alpha$. We denote this as $p^\alpha$.
We define the set of $\alpha$-bound plans in $P$ as $P^\alpha$.
We say a plan $p$ is $\alpha$-bound optimal if $\forall p' \in P^\alpha. \textsc{Time}(p) \leq \textsc{Time}(p'^\alpha)$. 
\end{definition}

@@

As a reference implementation, \ProjectName{} uses a general-purpose imputation strategy
based on chained-equation decision trees.  Chained equation
imputation methods~\cite{vanbuuren2011mice} (sometimes called \textit{iterative
regression}~\cite{gelman2006data}) impute multiple missing attributes by iteratively fitting
a predictive model of one missing attribute, conditional on all complete attributes
and all other missing attributes. Chained-equation decision tree algorithms are found to be
effective~\cite{akande2015empirical} and are widely used in epidemiological
domains~\cite{burgette2010multiple}. For our experiments, we use Weka's {\tt REPTree}
implementation (\cite{witten2016data}).

>>>>

\subsection{Imputation operators}
\label{sec:operators}
We introduce two new relational operators to perform imputation: \textit{Impute} ($\mu$) and
\textit{Drop} ($\delta$). Each operator takes arguments $(C, R)$ where $C$ is a set of
attributes and $R$ is a relation. \textit{Impute} uses a machine learning algorithm to
replace all \nullv{} values with non-null values for attributes in $C$ in the relation $R$ (discussed in detail in~\Cref{sec:imputation}).
In our experiments, we implement \textit{Impute} using a chained-equation decision trees
algorithm, but we highlight that
our system is agnostic to this choice.
\textit{Drop} simply removes from $R$ all tuples which have a \nullv{} value for some attribute in $C$.
Both operators guarantee that the resulting relation will contain no \nullv{} values for
attributes in $C$.

@@

\begin{definition}
Let $P$ be a set of plans. For $p \in P$, we say $p$ is $\alpha$-bound, if $(\textsc{Loss}(p) - \min_{p' \in P}(\textsc{Loss}(p')) \leq \alpha$. We denote this as $p^\alpha$.
We define the set of $\alpha$-bound plans in $P$ as $P^\alpha$.
We say a plan $p^\alpha$ is $\alpha$-bound optimal if $\forall p' \in P^\alpha. \textsc{Time}(p) \leq \textsc{Time}(p')$. 
\end{definition}

@@

As a reference implementation, \ProjectName{} uses a general-purpose imputation strategy
based on chained-equation decision trees.  Chained equation
imputation methods~\cite{vanbuuren2011mice} (sometimes called \textit{iterative
regression}~\cite{gelman2006data}) impute multiple missing attributes by iteratively fitting
a predictive model of one missing attribute, conditional on all complete attributes
and all other missing attributes. Chained-equation decision trees algorithms are found to be
effective~\cite{akande2015empirical} and are widely used in epidemiological
domains~\cite{burgette2010multiple}. For our experiments, we use Weka's {\tt REPTree}
implementation (\cite{witten2016data}).
